Stack Safety as a Security Property

### 2021-03-01 main.tex
Cleaned up formalism

\begin{figure}[b]
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int x = 42;
  f();
  print(x);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int x = 1;
  f();
  if(x)
    g();
  else
    h();
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int x = 1;
  f();
  x = -x;
  f();
  print(x);
}
\end{verbatim}
  \end{minipage}
  \caption{Integrity Examples}
  \label{fig:intexamples}
\end{figure}

\begin{figure}[b]
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int key = get_key();
  f();
  use_key(key);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int x = get_secret();
  if(x)
    g();
  else
    h();
  f();
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int x = 0;
  int y = get_secret();
  f();
  print(x);
}
\end{verbatim}
  \end{minipage}
  \caption{Confidentiality Examples}
  \label{fig:confexamples}
\end{figure}

@@

When does the programmer rely on stack safety? The most natural example is the first
in figure \ref{fig:intexample}, in which {\tt main} calls stores 42 to a local variable {\tt x},
calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
overwrite it's caller's stack data. To this end, our properties support a
{\em call rule} that can be used to reason that on return from {\tt f}, {\tt x} still
contains the same value. This is not only useful for the programmer: consider the second
example, where the call to {\tt h} should be unreachable and therefore subject to dead code
elimination, if only the compiler can soundly reason that {\tt x} does not change.
The property that enables the call rule is, in security terms, {\em integrity}: callers'
data is protected from modification by callees.

Alongside integrity, we have {\em confidentiality}, in which callers' data is protected
from leakage by their callees. Consider the first example in figure \ref{fig:confexamples},
in which {\tt main} recieves a key, then calls {\tt f}. If {\tt f} is an untrusted function,
it would be bad if it could just read {\tt key}. But there are more subtle examples,
such as the second one, where a secret value controls which function of {\tt g} and
{\tt h} are called prior to calling {\tt f}. Here, if {\tt g} or {\tt h} leaves behind
identifying information in unused stack space, {\tt f} could still deduce information
about the secret. So our confidentiality properties also support a {\em insensitivity rule},
that each function's behavior is independent of the state of its caller, other than arguments.
This also makes it easier, though not trivial, to prove idempotence: the channels through
which a callee can remember its own history are limited to those outside the stack.

Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
callers from their callees. Consider the last example in figure \ref{fig:intexample},
a variant of the famous ``awkward example,'' in which {\tt f} is called twice and may not
be idempotent. It should be the case that {\tt main} prints -1, but if
{\tt f}'s first activation stores its return address somewhere accessible, the second
activation can return to before {\tt x} is negated, cause {\tt x} to be set to
1 again before a third activation that would return normally. Then {\tt main} would print 1.

Preventing this form of reentrancy attack is the intuition behind
\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
the call graph itself. A callee should pass control back to the caller only by
transferring control to the instruction following the call; it should not be allowed
to jump back into the caller's code at an arbitrary point, either directly (via a jump
instruction) or indirectly (by corrupting stack data and returning). It also maintains
the integrity of the stack structure; otherwise a callee could leak data in the third
example of \ref{fig:conf-example} by failing to properly restore the stack pointer.
As {\tt x} is accessed by its offset from the stack pointer, an adversary could leave
the stack pointer so that {\tt x}'s offset now points to {\tt y}, and trick {\tt main}
into printing its secret.

@@

A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
\(\TRACEELEMS\). We will mostly be concerned with sequences of pairs of machine states
and observations, called \emph{machine traces} and denoted \(\machT \in \MOTRACE\),
and with sequences of 4-tuples of a machine state, a policy state, a context, and an
observation, called \emph{MPC-traces} and denoted \(\MPCT \in \MPCOTRACE\).

@@

The ``trace-of'' operator, written \(\mach \hookrightarrow \machT\),
coinductively relates a machine state with the trace of machine states
and observations produced by repeated application of step. It's defined in
terms of a coinductive helper, \(\mathit{obsforward}\):

\judgmenttwo{\(\mach \stepstoobs[\obs'] \mach'\)}
            {\(\mathit{obsforward} ~ \mach' ~ \obs' = \machT\)}
            {\(\mathit{obsforward} ~ \mach ~ \obs = (\mach,\obs) \machT\)}%
%
\judgment{\(\mathit{obsforward} ~ \mach ~ \tau = \machT\)}
         {\(\mach \hookrightarrow \machT\)}
\noindent
(We use juxtaposition of an element and a sequence to represent ``cons.'')
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

In our running example in \Cref{fig:running-trace-b} we can
see a depiction of (a prefix of) such a trace: starting from an initial machine
state at the very top, each instruction leads to a new machine state
(the diff of which, compared to the previous one, appears in the
``{\textit{Effect}'' column), and the majority of observations are $\tau$, except
for instructions 31 and 13, which are writes to $\rout$
where the observations are $0$ and $60$, respectively.

@@

We will also use \(\pi_m\), \(\pi_p\), \(\pi_c\), and \(\pi_o\) to project the components out of an
MPC state, and overload them to project a trace of the components out of an MPC trace.

@@

Recall that our properties use a compiler-derived code-map determine
which addresses are instructions and which functions they belong to, as well
as annotating them as calls, returns, and yields, as well as sharing information.
We now extend it with an annotation for valid entry points:
\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap, \entrymap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

\definition%{Control Separation}
A system enjoys {\em control separation} when each reachable step \(\mpcstatename\)
where 
% with respect to a function map
%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if, for any initial state \((\mach_0, \pol_0)\) and any adjacent pair of states
\(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
in the induced trace from \(\mpstate{\mach_0}{\pol_0}\),
if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
then \(\mpstate{\mach_1}{\pol_1}\) is either a call \(\callmap(\mach_1(\PCname))\)
or a return \(\retmap(\mach_1(\PCname))\).

\smallskip

For example, consider changing line 5 of the running example to a $\jalr$,
$\PCname \leftarrow r_4$, with effect $\PCname \leftarrow 42$, causing {\tt main}
to jump into the middle of {\tt g}.
This violates control separation because the instruction that causes a
function switch is not a valid call (or a return) according to the
call map. On the other hand we do permit functions to jump freely internally.

\definition%{Entry Integrity}
A system enjoys {\em entry integrity}
%with respect to a call map
%\(\callmap\) and entry map \(\entmap\)
if, for any initial state \((\mach_0, \pol_0)\)
and any n-transition from \((\mach_c,\pol_c)\) to
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a valid entry:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

The changed program described just above also violates entry integrity,
because the target (instruction 42) of the new $\jalr$ is not an established
entry point according to the entry map. Our properties do not enforce which
call points jump to which entry points, however; that would require a stronger
notion of control flow integrity.

\definition%{Return Integrity}
A system enjoys {\em return integrity} if all of its returns match prior calls
and return to the appropriate location.
For any initial state and its induced trace \((\mach_0\, \pol_0) \hookrightarrow \MPT\), two
conditions must hold.
\emph{(i)} \(\MPT\) must not have an unmatched return, i.e., \(\Unmatched ~ \MPT ~ \MPT_{\mathit{tail}}\)
must not hold for any \(\MPT_{\mathit{tail}}\).
%
\emph{(ii)} For any state pair \((\mach_c, \pol_c)\)
corresponding to a call in \(\MPT\), if
\(\MPT_{\mathit{call}}\) is the suffix of the trace starting at the call, and
\(\mpstatename = \mpostate{\mach_r}{\_}{\_}\)
is the first unmatched return such that
\( \Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{tail}}\) and \(\head{(\MPT_{\mathit{tail}})} = \mpstatename\),
then $\mach_r$ must correspond to a return state for $\mach_c$, i.e.,
\(\ret{\mach_c}\ \mach_r\) holds.

For example, if we modified the running example by adding
an instruction $ \rra \rightarrow \rra - 19$
between 45 and 46, this would cause the return from {\tt g} to go to instruction 8 and
return to {\tt main} instead of {\tt f}. While the instruction that
causes the return would still be a valid return (satisfying control
separation), the state after the return would not correspond to an
unmatched return for {\tt f}'s call to {\tt g}, violating return
integrity.

@@

\subsection{A Conservative Policy}
\label{sec:conservative}
%
The first micropolicy we present to enforce the definition of stack
safety developed in \cref{sec:lse-and-wbcf} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\subsection{A Stronger Lockstep Stack Safety Property}
\label{sec:lockstep}

%% \leo{Maybe lockstep?}

This conservative policy satisfies not only the stack safety policy of
\cref{sec:lse-and-wbcf},
but also a provably (in Coq) stronger one that enforces local state
encapsulation at every point during execution. To encode that
property, we not only keep track of the execution of one machine
(whose trace we are trying to characterize as safe), but of an entire
stack of machine variations corresponding to nested calls of the
original machine. For each step of the original machine, we take a
step in every variant machine (thus, keeping them in lockstep
execution) and enforce a single-step variant of integrity and
confidentiality tests. In addition, every time a call is made we push
a variation of the original machine state at the time of that call on
top of the variant stack (just like the stack-confidentiality property
enforced trace confidentiality for every call transition), and every time
the machine returns from a call we pop the top variation (to account
for stack confidentiality taking the prefix of the transition trace
until its return).

We start by defining the integrity and confidentiality tests, as well
as the variant stack, before giving the full definition of lockstep
stack safety.

\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
\newcommand*{\testInt}[3]{\ensuremath{\mathit{int}_T ~ #1 ~ #2 ~ #3}}
\newcommand*{\testConf}[2]{\ensuremath{\mathit{conf}_T ~ #1 ~ #2}}
\newcommand*{\testVS}[2]{\ensuremath{\mathit{\vs}_T ~ #1 ~ #2}}
\newcommand*{\testProp}[3]{\ensuremath{\mathit{stack}_T ~ #1 ~ #2 ~ #3}}

\definition
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \testInt{\contour}{\mach}{\mach'} \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
    \]
    \ifaftersubmission
\apt{We could have used this to define trace integrity in the first place...}
\fi
\definition Two machine states $\mach_1$ and $\mach_2$ satisfy the
\emph{stack confidentiality test} if the machines they stepped to
have their changed components coincide and the observations during
these steps are equal:
\[
\begin{array}{rl}
  \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \triangleq &
  \exists \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2'. ~
  \mach_1 \stepstoobs[\obs_1] \mach_1' \wedge \mach_2 \stepstoobs[\obs_2] \mach_2' \Rightarrow \\
&    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
\end{array}
\]
\ifaftersubmission
\apt{And the last clause here could have been used for trace confidentiality...}
\fi

\definition The testing property keeps track of a variation machine
for every nested call performed so far on a {\em variant stack}: a
list of machine states, each annotated with its own contour (written
as $\vm{C}$), and each being a variation of the actual machine
according to $C$.%
%
\footnote{The full formal definition in Coq also records additional
  information (the initial machine state at the call, its initial
  variation, the contour, and a return predicate) which we elide
  here for simplicity of exposition.}
%%
%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
%  $$\vs \in \VSS$$
%%
%\iflater
%\leo{This can be elided}
%  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
%  respect to it if the stack is nonempty, the return map of the bottom element
%  is the false property, and for each element $\vse \in \vs$:
%  \begin{itemize}
%  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
%  \item $\mach$ is in the trace $\machT$ of its initial machine,
%    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
%  \item Its variant state $\mathit{curr\_variant}(\vse)$
%    is in the trace of its initial variant,
%    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
%  \item The return function is in the return map if not the last element. \rb{TODO}
%  \item The registers of its contour $\contour(\vse)$ are public.
%  \end{itemize}
%  \fi

%% \definition
%% %  Let $\callmap$ a call map and $\codemap$ a function map.
%%   Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
%%   variant stack $\vs$, we say that an MP-trace induced from $(\mach,
%%   \pol)$ satisfies the \emph{strong stack safety} at a given call
%%   depth as follows.
%% % Refactor repetitive notation before case split
%%   In all cases that let $(\mach, \pol) \hookrightarrow_P
%%   (\mach', \pol', \obs)$.
%%   %and assume (unless the step is a return) that the
%%   %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
%%   %machines can step.

%% %  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
%% %  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
%% %  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
%% %  testing property} is satisfied at call depth $d$, and write
%% %  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:

%%   \begin{itemize}

%%   \item
%% %
%%     If the machine takes a \emph{call step} and $\contour$ is the contour at
%%     the call then,
%% %
%%     (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
%%     satisfied,
%% %
%%     (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
%%     satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
%%     is satisfied with resepct to an arbitrary $\contour$-variant
%%     of $\mach$,
%% =======
We also need two auxiliary notations that relate to variant stacks.
%
First, lockstep stack safety will check that the basic tests above hold for every
machine in the variant stack.  Given a machine that steps $(\mach,
\pol) \stepstopol[\obs] (\mach', \pol')$, say that a variant
stack $\vs$ satisfies the \emph{variant stack test} if for each
variant element $\vm{C}$ (which is a variation of $\mach$ for $C$),
$\vm{C}$ can take a step, denoted \(\vm{C} \stepstoobs[\obs] \mach'\),
while satisfying the stack integrity test \(\testInt{C}{\mach}{\mach'}\)
and stack confidentiality test $\testConf{\mach}{\mach'}$ with regard
to the original machine state. We will write $\testVS{(\mach,
  \pol)}{\vs}$ for this compound check.
%\apt{I'm having trouble here and below with
%  the scoping of $C$. Each variant stack is for one particular $C$, right?
%  So why not write $vs_C$ rather than $vm_C$ ?}
% LEO: Each machine has its own contour. Clarified above, hopefully
%% >>>>>>> 1886f7a7bf858ce0c9518e407d9481518a4f028d
%
Second, we will write $\vs \stepsto \vs'$ if every variant machine
in $\vs$ steps to the corresponding machine in $\vs'$.
%

We are now ready to define lockstep stack safety coinductively.

\definition Given a state $(\mach, \pol) \in \MPS$ and a variant stack
$\vs$ in which every element $\vm{\contour}$ is a variation of $\mach$
with respect to its annotated contour $\contour$, we say that an
MP-trace induced from $(\mach, \pol)$ satisfies the \emph{trace
  lockstep stack safety}, written $\testProp{}{\MPT}{\vs}$, if the
following conditions are satisfied:

\item
  If the machine takes a \emph{non-call, non-return step}
  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)\MPT,\]
  then the variant stack test must be satisfied
  \[\testVS{(\mach, \pol)}{\vs},\]
  and trace lockstep stack safety must be coinductively satisfied for
  the remainder of the trace and the variant stack $vs'$ that $vs$
  steps to ($\vs \stepsto \vs'$):
  \[ \testProp{}{((\mach',\pol',\obs)\MPT)}{\vs'}\]

\item
  If the machine takes a \emph{call step} and $\contour$ is the contour at
  the call
  %% \rb{Corrected the step function here, or should it be
  %% $\hookrightarrow$ and give the rest of the induced trace?}
  \[(\mach, \pol) \stepstopol[\obs] (\mach', \pol') \wedge \callmap(\mach) = n,\]
%
  then the variant stack test must be
  satisfied
  \[\testVS{(\mach, \pol)}{\vs},\]
%
%    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
%    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
%    is satisfied with regard to an arbitrary $\contour$-variant
%    of $\mach$,
%
  and trace lockstep stack safety must be coinductively satisfied
  for the remainder of the trace and for every variant stack
  that is the result of adding a variation of $\mach'$ with
  respect to $C$ at the top of the stack $\vs'$ that $\vs$ steps
  to:
  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\vm{\contour} : \vs')}\]

\item
%
  If the machine $\mach$ takes a \emph{return step},
  \[(\mach, \pol) \stepstopol[\obs] (\mach', \pol') \wedge \retmap(\mach),\]
  %
  then the top of the variant stack also returns,
  the tail of $\vs$ satisfies the stack safety test
  \[\testVS{(\mach, \pol)}{\mathit{tail}(\vs)},\]
  %
  and trace lockstep stack safety must be coinductively satisfied for
  the remainder of the trace and for the tail of the variant stack
  that $\vs$ steps to:
  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\mathit{tail}(\vs'))}\]

\item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
    (\mach, \pol, \tau)$, the property is satisfied for the singleton trace:
    $\testProp{}{[(\mach, \pol)]}{\vs}$.

%  \item If the machine takes a \emph{simultaneous return-call step}, that is,
%    the instruction to which execution returns immediately after completing a
%    call is itself a call instruction, the \emph{call step} and \emph{return
%    step} cases are combined.

\definition
%
A system satisfies {\em lockstep stack safety} if, for every initial machine $(\mach,\pol)$,
the trace induced from $(\mach, \pol)$ satisfies trace lockstep stack safety with
respect to the empty stack.

\medskip
\textsc{Theorem.}
Lockstep stack safety implies stack safety.
\smallskip

The proof is a fairly involved coinductive argument
(around 3K LoC in Coq) on the definition of trace lockstep stack safety,
using an auxiliary depth-index to ensure the proper alignment of
calls and returns.

>>>>

\begin{figure}[b]
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 42;
  f();
  print(x);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 1;
  f();
  if(x)
    g();
  else
    h();
}
\end{verbatim}
    \end{minipage}
  \caption{Simple Examples}
  \label{fig:examples1}
\end{figure}

\begin{figure}[b]
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main(int arg) {
  if(arg)
    g();
  else
    h();
  f();
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 1;
  f();
  x = -x;
  f();
  print(x);
}
\end{verbatim}
  \end{minipage}

  \caption{More Examples}
  \label{fig:examples2}
\end{figure}

@@

When does the programmer rely on stack safety? The most natural example is the first
in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
overwrite it's caller's stack data. To this end, our properties support a
{\em call rule} that can be used to reason that on return from {\tt f}, {\tt x} still
contains the same value. This is not only useful for the programmer: consider the second
example, where the call to {\tt h} should be unreachable and therefore subject to dead code
elimination, if only the compiler can soundly reason that {\tt x} does not change.
The property that enables the call rule is, in security terms, {\em integrity}: callers'
data is protected from modification by callees.

Alongside integrity, we have {\em confidentiality}, in which callers' data is protected
from leakage by their callees. Consider again the first example in figure \ref{fig:examples1},
in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
programmer had initialized {\tt x} after calling {\tt f} rather than before; such a change
should not alter the behavior of {\tt main}, but if {\tt f} can directly read {\tt x}, it
may. But there are more subtle examples, such as the first in figure \ref{fig:examples2},
where {\tt main} may call {\tt g} or {\tt h} are called prior to calling {\tt f}, depending on
its argument. Here, if {\tt g} or {\tt h} leaves behind identifying information in unused stack
space, {\tt f} could still deduce information about the argument and behave differently, violating
the abstraction of its call. So our confidentiality properties also support a {\em insensitivity rule},
that each function's behavior is independent of the state of its caller, other than arguments.
This also makes it easier, though not trivial, to prove a function's idempotence: the channels through
which a callee can remember its own history are limited to those outside the stack.

Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
callers from their callees. Consider the last example in figure \ref{fig:examples2},
a variant of the famous ``awkward example,'' in which {\tt f} is called twice and may not
be idempotent. It should be the case that {\tt main} prints -1, but if
{\tt f}'s first activation stores its return address somewhere accessible, the second
activation can return to before {\tt x} is negated, causing {\tt x} to be set to
1 again before a third activation that would return normally. Then {\tt main} would print 1.
Once again this is rather unexpected behavior.

Preventing this form of reentrancy attack is the intuition behind
\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
the call graph itself. A callee should pass control back to the caller only by
transferring control to the instruction following the call; it should not be allowed
to jump back into the caller's code at an arbitrary point, either directly (via a jump
instruction) or indirectly (by corrupting stack data and returning).

@@

A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\), and of observations.

@@

The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
coinductively relates an initial state with the trace of states
produced by repeated application of step, and potentially cut short by
the policy component.

\paragraph*{Projections}

We project out the machine state of an MPC-state with \(\pi_\mach\), the
policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
these operators to project out of an machine-trace a trace of machine states,
policy states, and contexts, respectively.

\paragraph*{Observations-Of}

The ``observations-of'' operator, written \(\pi_\obs\), takes a machine trace and
creates a trace of observations corresponding to the observations of each step in
the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
initial observation, a machine-trace, and an observation trace:

\judgmentthree{\(\pi_\mach ~ \mpcstatename \stepstoobs[\obs'] \pi_\mach ~ \mpcstatename'\)}
              {\(\mpcstatename' \hookrightarrow \MPCT\)}
              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
\judgment{}
         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}

Then to take the observations from any particular trace, we start with a \(\tau\) observation:

\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
         {\(\pi_\obs ~ \MPCT = \obsT\)}
\noindent
(We use juxtaposition of an element and a sequence to represent ``cons.'')

@@

@@

Recall that our properties use a compiler-derived code-map determine
which addresses are instructions and which functions they belong to, as well
as annotating them as calls, returns, and yields, as well as sharing information.
We now extend it with an annotation for valid entry points:
\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

\definition%{Control Separation}

Let \(\mpcstatename\) be a reachable state whose program counter points to code in
function \(\funid\), and \(\mpcstatename \stepstopol \mpcstatename'\) where
\(\mpcstatename'\) is state whose program counter points to a different function
\(\funid'\). A system enjoys {\em control separation} when all such \(\mpcstatename\)
are annotated \(\callmap\), \(\retmap\), or \(\yieldmap\).

\definition%{Entry Integrity}
A system enjoys {\em entry integrity} when, for any reachable \(\mpcstatename\) and
\(\mpcstatename'\) such that \(\mpcstatename\) is annotated \(\callmap\) and
\(\mpcstatename \stepstopol \mpcstatename'\), \(\mpcstatename'\) is annotated \(\entmap\).

\definition%{Return Integrity}
A system enjoys {\em return integrity}... [this one needs to talk about context.]

\definition%{Yield-back Integrity}
A system enjoys {\em yield-back integrity}... [same.]

@@

\subsection{A Conservative Tag Policy}
\label{sec:conservative}
%
The first micropolicy we present to enforce the definition of stack
safety developed in \cref{sec:lse-and-wbcf} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\section{Extension to Cheri}

\section{Extension to Software-based Enforcement}
